---
title: "244 Lab 1"
author: "Lucas Boyd"
date: "1/6/2022"
output: html_document
---

```{r setup, includev= TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
### update.packages(ask = FALSE) - updates all packages

```

### Read in the data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```


### Wrangling and ggplot review

**Example 1:** Find counts of observation by legal status and do some wrangling. 

```{r}
# groupby and summarize

sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())

# different way plus a few new functions
top_5_status <- sf_trees %>% 
  count(legal_status)  %>% # count observations grouping by legal status
  drop_na(legal_status) %>% # drop nas only from legal status column
  rename(tree_count = n) %>% # rename the tree_count column to n
  relocate(tree_count) %>% # relocate a column to the front of the dataframe
  slice_max(tree_count, n = 5) %>% # take the top 5 tree counts
  arrange(desc(tree_count)) # order the data frame my tree count, desc () function reverses order


```

Make a graph of the top 5 that we wrangled above

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  # fct_reorder to reorder legal status by tree count, put them in ascending order
  geom_col(fill = "darkgreen") +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() + # flip the axes
  theme_minimal()
```

**Example 2:** Only keep observations where legal status is "Permitted Site" and Caretaker is "MTA" and store as 'permitted_data_df'

```{r}
permitted_data_df <- sf_trees %>% 
  filter(legal_status == "Permitted Site", caretaker == "MTA") # or filter %in% c()
```

**Example 3:** Only keep blackwood acacia trees and then only keep columns legal_status, date, latitude, longitude and store as blackwood_acacia_df

```{r}
blackwood_acacia_df <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% #str_detect() keeps rows that finds those words within the observation, even if there are other words in that cell
  select(legal_status, date, lat = latitude, long = longitude) # changing the name of columns within select. cool! 
  
  ### make a little plot 
  
  ggplot(data = blackwood_acacia_df, aes(x = lon, y = lat) +
           geom_point(color = "darkgreen") 
```

**Example 4:** separate()

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = " :: ") %>% # separate into these two columns, separate by ::
```

**Example 5:** unite() - the opposite!

```{r}
ex_5 <- sf_trees %>% 
  unite("id_status", legal_status, tree_id, sep = " DOPE ") # bringing two columns together
```






